PeerMind

### 2016-08-06 paper.tex
Citations.
Cleaning.
Delegation.

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage[pdftex]{hyperref}
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}

@@

% End of preamble. Here it comes the document.
\begin{document}

@@

\todo[inline]{TODO}

@@

>>>>

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage[pdftex]{hyperref}
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}
\usepackage{pgf}
\usepackage{tikz}
\usepackage{tkz-graph}
\usepackage{mathtools}

@@

\GraphInit[vstyle=Dijkstra]
\tikzset{
  EdgeStyle/.append style={->,bend left}
}

\hyphenation{PeerMind}

\begin{document}

@@

%\todo[inline]{TODO}

@@

Terminology:
\begin{itemize}
\item community -- group which is making a decision using PeerMind
\item member -- user of the app who is participating in decision making
\end{itemize}

% decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)

% motion

\section{Related work}

\cite{andersen2008trust, dirnstorfer2010voting, ford2002delegative, rodriguez2007smartocracy, yamakawa2007toward}

\section{Features}

\subsection{Dynamic quroum}

\subsection{Score voting}

\subsection{Delegation}

Because not every member of the community has time or knowledge to vote all the time, each member can delegate
decisions to one or more other members of the community.
When a member does not vote themselves, information how those delegates voted is combined into a vote of
the non-voting member.
If a delegate has not voted themselves either, their vote is computed through delegation as well, recursivelly.

Each member $M$ can declare their delegates and ratios $r$ in which delegates' votes are combined. For example:

\begin{itemize}
\item Member $M_0$ decides not to delegate at all.
\item Member $M_1$ delegates to $M_0$ and $M_3$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectivelly.
\item Member $M_2$ delegates to $M_3$ with ratio $1.00$ ($r_{2,3}$).
\item Member $M_3$ delegates to $M_1$ and $M_2$ with ratios $0.40$ ($r_{3,1}$) and $0.40$ ($r_{3,2}$).
\end{itemize}

Observe that member $M_3$ decided to use only $0.8$ of their vote in the case they do not vote themselves.
The rest ($0.2$) counts towards non-voting population.
In this way a member can decide to participate in voting through delegation, but not fully.
Member $M_0$ deciced to not participate at all and if they do not vote they are counted towards non-voting population.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \SetGraphUnit{2.5}

    \Vertex[Math,L=M_0]{M0}
    \EA[Math,L=M_1](M0){M1}
    \SO[Math,L=M_3](M1){M3}
    \WE[Math,L=M_2](M3){M2}

    \Edge[label=$0.40$](M1)(M3)
    \Edge[label=$0.75$](M0)(M1)
    \Edge[label=$0.25$](M3)(M1)

    \tikzset{EdgeStyle/.append style={bend right}}
    \Edge[label=$1.00$](M3)(M2)
    \Edge[label=$0.40$](M2)(M3)
  \end{tikzpicture}
  \caption{Example delegation network. Edges are directed towards a member who is delegating, representing how their
  vote is computed from delegates' votes. E.g., $V_{M_1} = 0.75 V_{M_0} + 0.25 V_{M_3}$}~\label{fig:delegation-network}
\end{figure}

We can represent these delegations as a graph we call \emph{delegation network} as seen in
Figure~\ref{fig:delegation-network}.
We direct edges towards a member who is delegating, to represent how their vote is computed from delegates' votes.
For example, member $M_1$ delegates the decision to members $M_0$ and $M_3$.
$M_0$ and $M_3$ become $M_1$'s delegates.
As a consequence, $M_1$'s vote ($V_{M_1}$) is computed from $M_0$'s and $M_3$'s votes ($V_{M_0}$, $V_{M_3}$).

A delegation network can be described as a system of equations.

\begin{displaymath}
\begin{array}{rcl}
V_{M_0} & = & 0 \\
V_{M_1} & = & r_{1,0} V_{M_0} + r_{1,3} V_{M_3} = 0.75 V_{M_0} + 0.25 V_{M_3} \\
V_{M_2} & = & r_{2,3} V_{M_3} = 1.00 V_{M_3} \\
V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} = 0.40 V_{M_1} + 0.40 V_{M_2} \\
\end{array}
\end{displaymath}

When only some members vote ($\mathbf{v}$) we can use a delegation network to compute votes
of all members ($\mathbf{x}$).
We convert a delegation network to a matrix $\mathrm{D}$:

\begin{displaymath}
\mathrm{D} = \left[d_{ij}\right] = \begin{dcases*}
 1 & if $i = j$ \\
 -r_{i,j} & \parbox[t]{.55\columnwidth}{if $M_i$ did not vote, but delegates to $M_j$ with ratio $r_{i,j}$} \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

We solve the matrix equation:

\begin{displaymath}
\mathrm{D} \mathbf{x_v} = \mathbf{v}
\end{displaymath}

To continue with our example, let's say that $M_0$ and $M_2$ voted for a decision with $0.7$ and against with $0.4$,
respectivelly.
We are using score voting on $[-1, 1]$ range, so those votes are represented with $V_{M_0} = 0.7$ and
$V_{M_2} = -0.4$.

\begin{displaymath}
\mathrm{D} = \left[ \begin{array}{cccc}
1 & 0 & 0 & 0 \\
-0.75 & 1 & -0.25 & 0 \\
0 & 0 & 1 & 0 \\
0 & -0.40 & -0.40 & 1 \\
\end{array} \right]
\end{displaymath}

\begin{displaymath}
\mathbf{v} = \left[ \begin{array}{c}
0.7 \\
0 \\
-0.4 \\
0 \\
\end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
0.7 \\
0.54 \\
-0.4 \\
0.056 \\
\end{array} \right]
\end{displaymath}

The result (average of scores) without delegation is $\overline{v} = 0.15$ and
with delegation is $\overline{x} = 0.22$.
Delegation pushes decisision slightly in favor.

There are few details we have to look into.
Despite delegation some votes or parts of votes can still happen to be discarded and should count
towards non-voting population.
Members might not declare their delegates, or they might declare them without using their vote fully.
Other members can delegate to them as well, potentially propagating lack of a vote further.
For example, if member $M_0$ would not vote, then their vote would be discarded.
Moreover, we could not use $M_0$'s vote to compute $M_1$'s vote fully, which would influence $M_3$'s vote as well.

One option is that we would normalize ratio in such case, so if $M_0$ does not vote, then $r_{1,3}$ is normalized
from $0.25$ to $1.0$.
We could do this automatically for all members, or we could even leave to members to configure
what to do in such case for them.
In current version of PeerMind we decided to not handle this situation in any special way and we leave
discarded votes to propagate.

As a consequence we have to determine how many votes have been discarded to know the size of the non-voting
population.
Another issue is also a question of abstentions.
In PeerMind members can abstain from voting which is different from simply not participating.
Abstentions are counted towards quorum, but they do not contribute to the result.
To address this, we define two vectors:

\begin{displaymath}
\mathrm{w} = \left[w_{i}\right] = \begin{dcases*}
 1 & if $M_i$ voted \\
 0 & otherwise \\
\end{dcases*},\quad \mathrm{a} = \left[a_{i}\right] = \begin{dcases*}
 1 & if $M_i$ abstained \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

Solving $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ gives
us number of votes and number of abstentions, respectivelly, by summing together all values in the
resulting vectors: $w = \sum \mathbf{x_w}$, $a = \sum \mathbf{x_a}$.

\todo[inline]{Is the result then $\frac{\sum \mathbf{x_v}}{w}$ or $\frac{\sum \mathbf{x_v}}{\textrm{number of entries
in}~\mathbf{x_w}~\textrm{which are non-zero}}$?}

We pass computed votes and number of votes and abstaintions into computation of dynamic quorum to determine
a quorum and confidence, in the same manner as if members voted themselves.

% In general, there are various ways to use this delegation scheme.
To whom to delegate could be done for each vote separately, or decisions could be categorized into various
categories and then delegations for that category would be used.
But for simplicity, PeerMind currently supports only defining delegates separately from a decision making process,
and those delegates are then in effect for all votes where a member does not vote themselves.
A member is invited to define initial delegations at a registration time, and can change them later on at will,
but independently from a particular decision being made.

\subsection{Integration with offline meetings}
